Effect of quality-adjusted production function inputs

### 2015-06-02 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Now calculate deltas on mse and lambda. This is for non-dynamic SR.

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  coeffs <- naturalCoef(mod)
  res <- rbind(list(Country=Country, flavor=flavor, energy=energy, model=model, nest=nest))
  res <- cbind(res, coeffs)
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("unadjusted", "adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("noE", "withE"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
@

@@

lag <- 1 # year
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id, lag=1, pad="tail"){
  df <- dynamicSR(model, id=id, lag=lag, pad=pad)
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  df$Year <- model$data[["Year"]]
  return(df)  
}) %>% do.call(plyr::rbind.fill, .)

@@

Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
of the Solow residual~($\lambda$) over time
as calculated by Equation~\ref{eq:lambdaij} 
with the timespan for finite differences ($t_j - t_i$) set to 
\Sexpr{lag}~\Sexpr{if (lag == 1) "year" else "years"}.
It appears that quality-adjusting the factors of production
\emph{and} adding augmenting the production function with energy
decreases the Solow residual~($\lambda$),
i.e., brings the Solow residual closer to zero.

>>>>

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  res <- rbind(list(Country=Country, flavor=flavor, energy=energy, model=model, nest=nest))
  coeffs <- naturalCoef(mod)
  res <- cbind(res, coeffs)
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(mod)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
# Get the unadjusted, no-energy values for mse and lambda and place them on each appropriate row
AllCoeffs_unadj_noE <- filter(AllCoeffs, flavor=="unadjusted", energy=="noE") %>% 
                              select(Country, flavor, energy, model, nest, mse, lambda)
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="mse"] <- "mse_unadj_noE"
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="lambda"] <- "lambda_unadj_noE"
AllCoeffs <- merge(AllCoeffs, AllCoeffs_unadj_noE[c("Country", "model", "mse_unadj_noE", "lambda_unadj_noE")])
# Calculate differences relative to the unadjusted, no-energy condition
AllCoeffs <- mutate(AllCoeffs, Dlambda = abs(lambda) - abs(lambda_unadj_noE), Dmse = mse - mse_unadj_noE)

# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("unadjusted", "adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("noE", "withE"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
# Rearrange for easier viewing
AllCoeffs <- arrange(AllCoeffs, Country, model, flavor, energy, nest)
@

@@

lag.to.use <- 1 # years
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), 
                        f = function(model, id, lag=lag.to.use, pad="tail"){
print(lag)
  df <- dynamicSR(model, id=id, lag=lag, pad=pad)
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  df$Year <- model$data[["Year"]]
  return(df)  
}) %>% do.call(plyr::rbind.fill, .)

@@

Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
of the Solow residual~($\lambda$) over time
as calculated by Equation~\ref{eq:lambdaij} 
with the timespan for finite differences ($t_j - t_i$) set to 
\Sexpr{lag.to.use}~\Sexpr{if (lag.to.use == 1) "year" else "years"}.
It appears that quality-adjusting the factors of production
\emph{and} adding augmenting the production function with energy
decreases the Solow residual~($\lambda$),
i.e., brings the Solow residual closer to zero.
